Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-09-20 ICLR25/Paper_noProof.tex
exp + conclu

In this paper, we introduced the notion of {\em compensating pairs of paths}, with rationale why such a phenomenon creates inaccuracies hard to handle when verifying DNNs. We proved that this phenomenon is actually explaining entirely the inaccuracies, as in their absence, even the simplest Box abstraction (interval arithmetic) suffices to verify accurately DNNs. This is experimentally confirmed by the fact that DNNs harder to verify (verification-agnostic) also exhibit paths with larger compensating strength than  robustly-trained DNNs.

Based on this idea of compensating pairs of paths, we proposed the $\MILP_{Z}$ abstraction considering a subset $Z$ of the set of unstable ReLU nodes, that we proved to be fully accurate if {\em all} the compensating paths are covered by $Z$. Our empirical studies revealed that CMP, selecting $Z$ to cover only {\em a few} paths with the most significant compensation, can yield highly accurate results, up to 20\% more accurate than SOTA within the same runtime, again validating the focus on compensating strength. 
This underscores the potential of compensating strength as an innovative and promising avenue for enhancing DNN verification. We finally proposed several ways to use the compensating strength to optimize current tools in different directions, and leave that as future work.
\newpage

>>>>

In this paper, we studied the LP relaxation of the exact MILP encoding of ReLU-DNNs.
It allowed us to accurately assess when the relaxation of a node impacts very little 
the accuracy, and when it impacts the accuracy importantly, in which case representing it with a binary/integer variable is preferable. This allows us to select a small set of important neurons to encode as a binary variables, while improving accuracy drastically, necessitating $>4$x less integer variables than previous proposals \cite{DivideAndSlide} for the same accuracy. Our empirical studies revealed that Hybrid MILP, can yield highly accurate results, verifying up to 40\% more images than the SOTA ($\alpha,\beta$-Crown, winner of the 4 last VNNComp) within the same runtime. This opens a lot of perspective, among which: verifying efficiently hard instances; certifying robustness of images for larger $\epsilon$ neighbourhoods, verifying global rather than local properties \cite{lipshitz}.
\newpage

### 2024-09-20 ICLR25/experiments.tex
exp + conclu

Namely, these are 6 {\em hard} ReLU-DNNs: 5 MNIST DNN from the ERAN GitHub (\url{https://github.com/eth-sri/eran}, the 4th to the 8th DNNs provided) as well as 1 CIFAR CNN from 
\cite{AdversarialTrainingAndProvableDefenses}, also see \cite{SDPFI}. 
For reference, we also report an easy but very large ResNet Network for CIFAR10, already tested with $\alpha,\beta$ CROWN. We report in Table \ref{table_hybrid} the $\%$ of undecided images, that is the $\%$ of images than can be neither falsified (by $\alpha,\beta$-CROWN) nor verified by the tested verifier, among the 100 first images for each MNIST or CIFAR10 benchmark.

@@

>>>>

Namely, these are 6 {\em hard} ReLU-DNNs: 5 MNIST DNN from the ERAN GitHub (\url{https://github.com/eth-sri/eran}, the 4th to the 8th DNNs provided) as well as 1 CIFAR CNN from 
\cite{AdversarialTrainingAndProvableDefenses}, also see \cite{SDPFI}, with the same $\epsilon$ setting. For reference, we also report an easy but very large ResNet Network for CIFAR10, already tested with $\alpha,\beta$ CROWN. We report in Table \ref{table_hybrid} the $\%$ of undecided images, that is the $\%$ of images than can be neither falsified (by $\alpha,\beta$-CROWN) nor verified by the tested verifier, among the 100 first images for each MNIST or CIFAR10 benchmark, the exact same setting and DNNs as used in Table \ref{table_beta}.

@@

\subsection{Finer Grain evaluation of Accuracy}

In order to evaluate the accuracy of Hybrid MILP in a finer way, showcasing 
the capabilities to have a very accurate and efficient verifier, we consider
a quantitative questions for each image, rather than a pure yes (verified)/ no (not verified) question. Namely, we compute the $\epsilon$ for which the verifier can certify local-robustness around a given image, which makes sense as there is little rationale in setting a particular $\epsilon$ (which is however the usual local-robustness setting).

For that, we considered the DNN 6x500, and the 20 first images from the MNIST benchmark.
We first run the attack from $\alpha,\beta$-Crown for varying $\epsilon$, using a binary search, to set up an upper bound on $\epsilon$ (the initialization is $\epsilon \in [0,1]$).
Then we run binary search with a global time out of $10 000s$, initialized from $0$ to this upper bound, where each call is either to $\alpha,\beta$-Crown with TO=2000s, or Hybrid MILP.
We report the results (upper bound, best bound verified by $\alpha,\beta$-Crown and by Hybrid MILP) for each image. We report the results in Fig. \ref{fig2}

\begin{figure}[t!]
	\begin{centering}
\includegraphics[scale=0.5]{epsilon.png}.
\caption{$\epsilon$-robustness proved after $10 000s$ for each of the 20 first images of MNIST.}
\label{fig2}
	\end{centering}
\end{figure}

Analysis: in 90\% of the cases, Hybrid MILP is very close to the upper bound. On 2 images (10 and 13), Hybrid MILP is far from the upper bound: these are also the ones where the upper bound is the highest, so it is possible the upper bound did not find a close attack to robustness.

Compared with $\alpha,\beta$-CROWN, Hybrid MILP is much closer to the upper bound, but for 2 cases where $\alpha,\beta$-CROWN is already very close to the upper bond. On average, the $\epsilon$ gap to upper bound is $0.014$ for Hybrid MILP, and $0.42$ for 
$\alpha,\beta$-CROWN.

\subsection{Comparison with other Verifiers?}

We volontarily limited the comparison to $\alpha,\beta$-Crown because it is the most efficient verifier available to date, and it enabled us to consider much finer parameter choice without too much cluter.

Notice that several other verifiers were already reported \cite{crown} on these same DNNs, with defavorable comparison towards $\alpha,\beta$-CROWN (PRIMA \cite{prima}, SDP-FO \cite{SDPFI}, etc). Further, we reported accuracy of complete verifiers, NNenum \cite{nnenum}, Marabou \cite{Marabou, Marabou2}, respectively 4th, 3rd 
of the last VNNcomp'24 \cite{VNNcomp24}, as well as full MILP \cite{MILP} in Table \ref{table_complete}, showing that these verifiers are not competitive on hard instances either, even on the smallest hard DNN. Concerning MnBAB \cite{ferrari2022complete}, it compares slightly defavorably in time and accuracy towards $\alpha,\beta$-CROWN on CNN-B-Adv \cite{ferrari2022complete}, and we also run it for the MNIST DNNs, and it was slightly slower and less accurate than $\alpha,\beta$-CROWN at several time-out settings. 
Last, Pyrat code is not open source, which made running it impossible.
